---
title: "p8105_hw1_waa2119"
author: "William Anderson"
date: "2022-09-20"
output: github_document
---
```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```


#P8105 HW 1 Assignment

This loads the penguins data set
```{r setup}

data("penguins", package = "palmerpenguins")

```

* The penguins data set contains the following information in each column: `r colnames(penguins)`


  * The penguin species ranges from `r penguins[1, 1]`, `r penguins[190, 1]`, and `r penguins[340, 1]`
  

  * The islands these penguins reside on is `r penguins[1, 2]`, `r penguins[21, 2]` and `r penguins[36, 2]` 


  * The sex of these penguins are both `r penguins[1, 7]` and `r penguins[2, 7]`


  * The year this data was collected ranges between `r penguins[172, 8]`, `r penguins[66, 8]`, and `r penguins[109, 8]`


* The size of the penguins data set is `r nrow(penguins)` rows by  `r ncol(penguins)` columns

* The mean flipper length of all penguins in the data set is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm

Now we will make a scatter plot of flipper length vs bill length
```{r, scatter plot of flipper length vs bill length}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

```

