---
title: "p8105_hw1_waa2119"
author: "William Anderson"
date: "2022-09-20"
output: github_document
---

```{r, message = FALSE}

library(tidyverse)

```


# P8105 HW 1 Assignment


This loads the penguins data set
```{r}

data("penguins", package = "palmerpenguins")

head(penguins, 20)

```


# Problem 1


* The penguins data set contains the following information in each column: `r colnames(penguins)`


* The penguin species ranges from `r penguins[1, 1]`, `r penguins[190, 1]`, and `r penguins[340, 1]`. 


* The islands these penguins reside on is `r penguins[1, 2]`, `r penguins[21, 2]` and `r penguins[36, 2]`.


* The sex of these penguins are both `r penguins[1, 7]` and `r penguins[2, 7]`. 


* The year this data was collected ranges between `r penguins[172, 8]`, `r penguins[66, 8]`, and `r penguins[109, 8]`. 


* The mean, median, and standard deviation of the bill length of these penguins is `r mean(penguins$bill_length_mm, na.rm = TRUE)`, `r median(penguins$bill_length_mm, na.rm = TRUE)`, and `r sd(penguins$bill_length_mm, na.rm = TRUE)` mm respectively.


* The mean, median, and standard deviation of the bill depth of these penguins is `r mean(penguins$bill_depth_mm, na.rm = TRUE)`, `r median(penguins$bill_depth_mm, na.rm = TRUE)`, and `r sd(penguins$bill_depth_mm, na.rm = TRUE)` mm respectively. 


* The mean, median, and standard deviation of the flipper length of these penguins is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`, `r median(penguins$flipper_length_mm, na.rm = TRUE)`, and `r sd(penguins$flipper_length_mm, na.rm = TRUE)` mm respectively. 


* Lastly, the the mean, median, and standard deviation of the body mass of these penguins is `r mean(penguins$body_mass_g, na.rm = TRUE)`, `r median(penguins$body_mass_g, na.rm = TRUE)`, and `r sd(penguins$body_mass_g, na.rm = TRUE)` g respectively. 



* The size of the penguins data set is `r nrow(penguins)` rows by  `r ncol(penguins)` columns


* The mean flipper length of all penguins in the data set is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm



Now we will make a scatter plot of flipper length vs bill length

```{r, warning = FALSE}


ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()


ggsave("scatter_plot_HW1.pdf")

```



# Problem 2

Now creating a tibble with various vectors inside

```{r}

my_data_frame = tibble(
  
  my_random_sample = rnorm(n = 10, mean = 0, sd = 1),
  
  log_vec = my_random_sample > 0,
  
  char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  
  fact_vec = factor(c("First", "Second", "Third", "First", "Second", "Third", "First", "Second", "Third", "First"))
  
)

```



Now attempting to take the mean of the vectors within the tibble

```{r}

avg_rand_samp = mean(pull(my_data_frame, my_random_sample))

avg_log_vec = mean(pull(my_data_frame, log_vec))

avg_char_vec = mean(pull(my_data_frame, char_vec))

avg_fact_vec = mean(pull(my_data_frame, fact_vec))


```


We see that the mean of the random sample is `r avg_rand_samp`, and the mean of the logical vector is: `r avg_log_vec`, because R assigns the number 1 to true values and 0 to false values, but we are unable to compute the mean of the character vector: `r avg_char_vec`, and mean of factor vector: `r avg_fact_vec` as they are not numerical values. 


We can then convert their values to the numerical type

```{r, changing to numeric type, eval = FALSE}

num_log_vec = as.numeric(pull(my_data_frame, log_vec))

num_char_vec = as.numeric(pull(my_data_frame, char_vec))

num_fact_vec = as.numeric(pull(my_data_frame, fact_vec))


```

This converts non-numeric values such as characters and factor levels to numerical data which can then be used in mathematical formulas for computations.

