---
title: "p8105_hw1_waa2119"
author: "William Anderson"
date: "2022-09-20"
output: github_document
---
```{r, message = FALSE}

library(tidyverse)

```


#P8105 HW 1 Assignment

This loads the penguins data set
```{r setup}

data("penguins", package = "palmerpenguins")

```

# Problem 1

* The penguins data set contains the following information in each column: `r colnames(penguins)`


  * The penguin species ranges from `r penguins[1, 1]`, `r penguins[190, 1]`, and `r penguins[340, 1]`
  

  * The islands these penguins reside on is `r penguins[1, 2]`, `r penguins[21, 2]` and `r penguins[36, 2]` 


  * The sex of these penguins are both `r penguins[1, 7]` and `r penguins[2, 7]`


  * The year this data was collected ranges between `r penguins[172, 8]`, `r penguins[66, 8]`, and `r penguins[109, 8]`


* The size of the penguins data set is `r nrow(penguins)` rows by  `r ncol(penguins)` columns

* The mean flipper length of all penguins in the data set is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm


Now we will make a scatter plot of flipper length vs bill length

```{r, scatter plot of flipper length vs bill length}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("scatter_plot_HW1.pdf")

```

# Problem 2

Now creating a tibble with various vectors inside

```{r}

my_data_frame = tibble(
  
  my_random_sample = rnorm(n = 10, mean = 0, sd = 1),
  
  log_vec = my_random_sample > 0,
  
  char_vec = character(length = 10),
  
  fact_vec = factor(char_vec, levels = 3)
  
)

```

Now attempting to take the mean of the vectors within the tibble

```{r}

avg_rand_samp = mean(pull(my_data_frame, my_random_sample))

avg_log_vec = mean(pull(my_data_frame, log_vec))

avg_char_vec = mean(pull(my_data_frame, char_vec))

avg_fact_vec = mean(pull(my_data_frame, fact_vec))


```

We see that the mean of the random sample is `r avg_rand_samp`, but we are unable to calculate the mean of the logical: `r avg_log_vec`, mean of character: `r avg_char_vec`, and mean of factor vector: `r avg_fact_vec` as they are not numerical values

We can then convert their values to the numerical type

```{r, changing to numeric type, eval = FALSE}

num_log_vec = as.numeric(pull(my_data_frame, log_vec))

num_char_vec = as.numeric(pull(my_data_frame, char_vec))

num_fact_vec = as.numeric(pull(my_data_frame, fact_vec))


```


